An in-depth study of Explainability and Adversarial Robustness for RNNs

### 2019-12-12 ex3/paper/paper.tex
Change bib style to IEEE

\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage[inline]{enumitem}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{subfig}

@@

\begin{table}
% Evaluated using
% ./python learn.py --dataroot flows.pickle --net runs/Oct26_00-03-50_gpu/lstm_module_1284.pth --function test --batchSize 512
% ./python learn.py --dataroot flows15.pickle --net runs/Oct28_15-41-46_gpu/lstm_module_997.pth --function test --batchSize 512
% on Dec. 4th
\caption{Detection performance on a packet and flow level.} \label{tab:performance_results}
\centering
\begin{tabular}{l r r r r} \toprule
& \multicolumn{2}{c}{CIC-IDS-17} & \multicolumn{2}{c}{UNSW-NB-15} \\
	&	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
Accuracy	&	0.991	&	0.997	&	0.995	&	0.983	\\	
Precision	&	0.970	&	0.997	&	0.834	&	0.786	\\	
Recall	&	0.978	&	0.991	&	0.876	&	0.726	\\	
F1	&	0.974	&	0.994	&	0.855	&	0.755	\\	
Youden	&	0.972	&	0.990	&	0.873	&	0.719	\\

@@

In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed. If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the \gls{pdp} for feature $i$ can be defined as
\begin{equation} \label{eq:pdp}
\text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
\end{equation}

To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
\begin{equation} \label{eq:pdp_conditional}
\text{\gls{pdp}}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
\end{equation}
%
%\begin{figure*}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
%
%\caption{Examples of conditional PD plots for source and destination port. We only analyze these features with the \gls{pdp} since it is not straightforward to define the \gls{pdp} for features that can vary for each packet in a flow. The y-axis shows the probability for malicious traffic.}
%\todo{Max: Remove these? Not really interesting or related to the rest of the paper. Refer to other paper where we talk about this already.}
%\label{fig:pdp}
%\end{figure*}
%
In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class. Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class. The results (omitted for brevity) show that for some attack classes the port numbers play an important role. When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers. A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur. For a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.

@@

\section{Defences}
We now propose two approaches to increase robustness of the \gls{rnn} against adversarial attacks.
\subsection{Reducing features}
The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
\begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
\item Leaving out all features that are manipulable: packet size and \gls{iat}
\item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
\end{itemize}

@@

\bibliographystyle{plain}
\bibliography{bibliography}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage[inline]{enumitem}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{subfig}

@@

\begin{table}
% Evaluated using
% ./python learn.py --dataroot flows.pickle --net runs/Oct26_00-03-50_gpu/lstm_module_1284.pth --function test --batchSize 512
% ./python learn.py --dataroot flows15.pickle --net runs/Oct28_15-41-46_gpu/lstm_module_997.pth --function test --batchSize 512
% on Dec. 4th
\caption{Detection performance on a packet and flow level.} \label{tab:performance_results}
\centering
\begin{tabular}{l r r r r} \toprule
& \multicolumn{2}{c}{CIC-IDS-2017} & \multicolumn{2}{c}{UNSW-NB15} \\
	&	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
Accuracy	&	0.991	&	0.997	&	0.995	&	0.983	\\	
Precision	&	0.970	&	0.997	&	0.834	&	0.786	\\	
Recall	&	0.978	&	0.991	&	0.876	&	0.726	\\	
F1	&	0.974	&	0.994	&	0.855	&	0.755	\\	
Youden	&	0.972	&	0.990	&	0.873	&	0.719	\\

@@

In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed~\cite{friedman_greedy_2001}. To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}. If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space, $f(X) \in [0,1]$ denotes the neural network's prediction and $c$ the attack class, the \gls{pdp} for feature $i$ can be defined as
% \begin{equation} \label{eq:pdp}
% \text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
% \end{equation}

% To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
\begin{equation} \label{eq:pdp_conditional}
\text{\gls{pdp}}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
\end{equation}
%
%\begin{figure*}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
%
%\caption{Examples of conditional PD plots for source and destination port. We only analyze these features with the \gls{pdp} since it is not straightforward to define the \gls{pdp} for features that can vary for each packet in a flow. The y-axis shows the probability for malicious traffic.}
%\todo{Max: Remove these? Not really interesting or related to the rest of the paper. Refer to other paper where we talk about this already.}
%\label{fig:pdp}
%\end{figure*}
%
In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class. Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class. The results (omitted for brevity) show that for some attack classes the port numbers play an important role. When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers. A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur. For a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.

@@

\section{Defences}
We now propose two approaches to increase robustness of the \gls{rnn} against adversarial attacks.
\subsection{Reducing features}
The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
\begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
\item Leaving out all manipulable features: packet size and \gls{iat}
\item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
\end{itemize}

@@

@@

\renewcommand*{\bibfont}{\small}
\bibliographystyle{ieeetr}
\bibliography{bibliography}
